---
title: "B.C. Labour Force Survey Highlights"
output: 
  flexdashboard::flex_dashboard:
    theme: yeti
    orientation: columns
    vertical_layout: fill
    source_code: https://github.com/bcgov/labour-force-survey-digest
---


<!--
Copyright 2020 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE, cache=TRUE}
#load libraries in setup.R
source("setup.R")

#source data
# source("get-data.R")
lfc_province_tidy <- readRDS("tmp/lfc_province_tidy.rds")
lfc_region_tidy <- readRDS("tmp/lfc_region_tidy.rds")
employment_by_class_tidy <- readRDS("tmp/employment_by_class_tidy.rds")

#report month
report_month <- month(Sys.Date() %m-% months(1),
        label = TRUE, abbr = FALSE)

#report year
report_year <- year(Sys.Date())
previous_year <- year(Sys.Date()) - 1

#common chart themes
theme_vert_bar <- theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "grey50"),
    axis.text = element_text(size = 11),
    # legend.position = "top",
    # legend.direction = "horizontal",
    legend.text = element_text(size = 11))

theme_horiz_bar <- theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.line = element_line(colour = "grey50"),
    axis.text = element_text(size = 11),
    # legend.position = "top",
    # legend.direction = "horizontal",
    legend.text = element_text(size = 11))

theme_facet_bar <- theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "grey50"),
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 9),
    strip.text = element_text(size = 11))

recent_month_change_employment <- lfc_province_tidy %>%
  filter(vector == "v2064701",
         date == max(date)) 

year_change_employment <- lfc_province_tidy %>%
  filter(vector == "v2064701",
         date %in% c(min(date), max(date))) %>% 
  mutate(
    year_change = value - lag(value),
    year_change_percent = value / lag(value) - 1
  ) %>% 
  filter(date == max(date))
```


Employment
======================================================================


Column {data-width=30}
-----------------------------------------------------------------------


### Reference Date {.value-box}

```{r}
#LFS month-year
valueBox(value = paste(report_month, report_year),
icon = "fa-calendar")
```

### Month Change in Employment {.value-box}

```{r}
#recent monthly change in employment
valueBox(value = recent_month_change_employment %>% 
  pull(month_change) %>%
  signs(format = comma, add_plusses = TRUE),
         icon = "fa-users-cog",
         caption = paste("Change in Employment for", report_month))
```

### % Month Change in Employment {.value-box}

```{r}
valueBox(value = recent_month_change_employment %>%
  pull(month_change_percent) %>%
  signs(format = percent, add_plusses = TRUE, accuracy = 0.1),
         icon = "fa-users-cog",
         caption = paste("% Change in Employment for", report_month))
```

### Year Change in Employment {.value-box}

```{r}
#year over year change employment
valueBox(value = year_change_employment %>% 
  pull(year_change) %>%
  signs(format = comma, ),
         icon = "fa-users-cog",
         caption = paste("Change in Employment Since", report_month, previous_year))
```

### % Year Change in Employment {.value-box}

```{r}
#year over year change employment
valueBox(value = year_change_employment %>% 
  pull(year_change_percent) %>%
  signs(format = percent, add_plusses = TRUE, accuracy = 0.1),
         icon = "fa-users-cog",
         caption = paste("% Change in Employment Since", report_month, previous_year))
```


Column {data-width=500}
-----------------------------------------------------------------------

### `r report_year` Monthly Change in Employment

```{r}
lfc_province_tidy %>% 
  filter(vector == "v2064701",
         date > "2019-12-01" ) %>%
  ggplot(aes(x = date, y = month_change)) +
  geom_col(fill = "#1d91c0", alpha = 0.6) +
  geom_text(aes(label = signs(month_change_percent,
                              format = percent,
                              add_plusses = TRUE,
                              accuracy = 0.1)),
            nudge_y = 13000, colour = "grey30", size = 4) +
  labs(x = NULL,
       y = NULL) +
  scale_y_continuous(labels = comma) +
  theme_minimal() +
  theme_vert_bar
```

### Change in Employment for `r report_month` by Province

```{r}
lim <- c(0,
         lfc_province_tidy %>%
  ungroup() %>% 
  filter(labour_force_characteristics == "Employment",
         sex == "Both sexes",
         age_group == "15 years and over",
         date == max(date),
         geo != "Canada") %>% 
  slice_max(month_change, n = 1) %>% 
  pull(month_change) + 50000
)

lfc_province_tidy %>% 
  filter(labour_force_characteristics == "Employment",
         sex == "Both sexes",
         age_group == "15 years and over",
         date > "2019-12-01" ) %>%
  filter(date == max(date),
         geo != "Canada") %>% 
  mutate(fill_col = ifelse(geo == "British Columbia", "bc", "other")) %>% 
  ggplot(aes(x = reorder(geo, -month_change), y = month_change)) +
  geom_col(aes(fill = fill_col), alpha = 0.6) +
  coord_flip() +
  geom_text(aes(label = signs(month_change_percent,
                              format = percent,
                              add_plusses = TRUE,
                              accuracy = 0.1)),
            nudge_y = 25000, colour = "grey30", size = 3.5) +
  labs(x = NULL,
       y = NULL) +
  scale_y_continuous(labels = comma,
                     expand = c(0, 0),
                     limits = lim) +
  scale_fill_manual(guide = FALSE,
                    values = c(bc = "#FDE725FF",
                               other = "#365C8DFF")) +
  theme_minimal() +
  theme_horiz_bar
```



Column {data-width=500}
-----------------------------------------------------------------------

### `r report_year` % Monthly Change in Employment by Age & Gender

```{r}
lfc_province_tidy %>% 
  filter(date > "2019-12-01",
         geo == "British Columbia",
         labour_force_characteristics == "Employment",
         age_group %in% c("15 to 24 years",
                          "25 to 54 years",
                          "55 years and over")) %>%
  mutate(sex = recode(sex, "Both sexes" = "All")) %>% 
  filter(sex != "All") %>% 
  ggplot(aes(x = date, y = month_change_percent)) +
  geom_col(fill = "#1d91c0", alpha = 0.6) +
  facet_grid(sex ~ age_group) +
  # geom_text(aes(label = signs(month_change_percent,
  #                             format = percent,
  #                             add_plusses = TRUE,
  #                             accuracy = 0.1)),
  #           nudge_y = 13000, colour = "grey30", size = 2) +
  labs(x = NULL,
       y = NULL) +
  scale_y_continuous(labels = percent) +
  theme_minimal() +
  theme_facet_bar
```

### Changes in Employment for `r report_month` by Economic Region 

```{r}
lfc_region_tidy %>%
  filter(
    geo != "British Columbia",
    labour_force_characteristics == "Employment",
    date == max(date)) %>% 
  mutate(geo_label = str_wrap(str_extract(geo, "^([^,])+"), width = 10)) %>%
  st_as_sf() %>%
  rmapshaper::ms_clip(bcmaps::bc_bound(class = "sf")) %>% 
  ggplot() +
  geom_sf(aes(fill = month_change), colour = "white") +
  geom_sf_text(aes(label = geo_label),
               size = 2.5,
               colour = "white") +
  coord_sf(datum = NA) +
  labs(x = NULL, y = NULL,
       caption = "(3-month-moving-avg, seasonally unadjusted)") +
  scale_fill_viridis(name = paste("Change in\nEmployment\nfor", report_month),
                     direction = -1,
                     label = scales::comma) +
  theme_minimal() +
  theme(legend.title = element_text(size = 10),
        plot.caption = element_text(size = 8,
                                    colour = "grey60",
                                    face = "italic",
                                    hjust = 1.3))
  
# plotly::ggplotly(p) 
```


Unemployment
======================================================================

Column
-----------------------------------------------------------------------

### Reference Date {.value-box}

```{r}
#LFS month-year
valueBox(value = paste(
  month(Sys.Date() %m-% months(1),
        label = TRUE, abbr = FALSE),
  year(Sys.Date())
),
icon = "fa-calendar")
```


```{r}
all_month_change_unemploymt <- get_cansim_vector_for_latest_periods(c("all_unemployment" = "v2064704"), periods = 13) %>%
  normalize_cansim_values() %>%
  clean_names() %>%
  mutate(
    month_change = value - lag(value),
    month_change_percent = value / lag(value) - 1
  ) 

#monthly value change in employment
recent_month_value_change_unemploymt <- all_month_change_unemploymt %>%
  filter(year(date) == year(Sys.Date()),
         month(date) == month(Sys.Date()) - 1) %>%
  pull(month_change) %>%
  signs(format = comma, add_plusses = TRUE)

#monthly % change in employment
recent_month_perc_change_unemploymt <- all_month_change_unemploymt %>%
  filter(year(date) == year(Sys.Date()),
         month(date) == month(Sys.Date()) - 1) %>%
  pull(month_change_percent) %>%
  signs(format = percent, add_plusses = TRUE, accuracy = 0.1)
```

### Month Change in Unemployment {.value-box}

```{r}
valueBox(value = recent_month_value_change_unemploymt,
         icon = "fa-users-cog")
```

### % Month Change in Unemployment {.value-box}

```{r}
valueBox(value = recent_month_perc_change_unemploymt,
         icon = "fa-users-cog")
```







About
======================================================================

Data are sourced from the [Statistics Canada Labour Force Survey (LFS)](https://www.statcan.gc.ca/eng/survey/household/3701), a monthly survey which measures the current state of the Canadian labour market. 
  
Statistics Canada data are released under the [Statistics Canada Open Licence](https://www.statcan.gc.ca/eng/reference/licence).
  
Statistics Canada Labour Force Survey Tables: 

 - [Statistics Canada Table **14-10-0287-03**: Labour force characteristics by province, monthly, seasonally adjusted](https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410028703)
 - [Statistics Canada Table **14-10-0293-02**: Labour force characteristics by economic region, three-month moving average, unadjusted for seasonality](https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410029302)
 - [Statistics Canada Table **14-10-0288-01**: Employment by class of worker, monthly, seasonally adjusted and unadjusted, last 5 months](https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410028801)
 - [Statistics Canada Table **14-10-0127-01**: Reason for not looking for work, monthly, unadjusted for seasonality](https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410012701&pickMembers%5B0%5D=1.11&pickMembers%5B1%5D=3.1&pickMembers%5B2%5D=4.1)
 
B.C. Data Catalogue:
 - [Current Census Economic Regions](https://catalogue.data.gov.bc.ca/dataset/1aebc451-a41c-496f-8b18-6f414cde93b7)

This report is built with the [`flexdashboard` R package](https://rmarkdown.rstudio.com/flexdashboard/), sourcing the Statistics Canada Labour Force Survey data using the [`cansim` R package](https://mountainmath.github.io/cansim/index.html), and the British Columbia economic geospatial data using the [`bcdata` R package](https://bcgov.github.io/bcdata/). 
  
The [R](https://www.r-project.org/) code is available in bcgov GitHub [here](https://github.com/bcgov/labour-force-survey-digest), available under the [Apache 2.0 licence](https://github.com/bcgov/labour-force-survey-digest/blob/master/LICENSE).
