---
title: "Labour Force Survey Highlights"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
geometry: margin = 1in
urlcolor: blue
fontsize: 11pt
header-includes:
- \usepackage[document]{ragged2e}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, include = FALSE)

library(cansim)
library(dplyr)
library(ggplot2)
library(janitor)
library(lubridate)
library(scales)

## Statistics Canada Table "14-10-0287-03"
## Labour force characteristics by province, monthly, seasonally adjusted
## https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410028703

## v2064701 (British Columbia, Employment, Both sexes, 15 years and over, Estimate, Seasonally adjusted)
## v2064728 (British Columbia, Employment, Both sexes, 15 to 24 years, Estimate, Seasonally adjusted)
## v2064836 (British Columbia, Employment, Both sexes, 25 to 54 years, Estimate, Seasonally adjusted)
## v2064818 (British Columbia, Employment, Males, 25 years and over, Estimate, Seasonally adjusted)
## v2064827 (British Columbia, Employment, Females, 25 years and over, Estimate, Seasonally adjusted)

vectors <- c(
  "all_employment" = "v2064701",
  "youth_employment" = "v2064728",
  "core_working_age_employment" = "v2064836",
  "males_25over_employment" = "v2064818",
  "females_25over_employment" = "v2064827"
)

df <- get_cansim_vector_for_latest_periods(vectors,
                                           periods = 13) %>%
  normalize_cansim_values() %>%
  clean_names()

df_month_over_month_changes <- df %>%
  group_by(vector) %>%
  mutate(
    month_change = value - lag(value),
    month_change_percent = round((value / lag(value) - 1) * 100,
                                 digits = 1)
  )

df_year_over_year_changes <- df %>%
  group_by(vector) %>%
  filter(month(date) == month(Sys.Date()) - 1) %>%
  mutate(
    year_change = value - lag(value),
    year_change_percent = round((value / lag(value) - 1) * 100,
                                 digits = 1)
  )
```

### Employment Highlights

• Following the decline in employment of `r df_month_over_month_changes %>% filter(vector == "v2064701", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 2) %>% pull(month_change) %>% comma()` (`r df_month_over_month_changes %>% filter(vector == "v2064701", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 2) %>% pull(month_change_percent)`%) in `r df_month_over_month_changes %>% filter(vector == "v2064701", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 2) %>% pull(date) %>% format("%B")`, there was an increase of `r df_month_over_month_changes %>% filter(vector == "v2064701", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()` (`r df_month_over_month_changes %>% filter(vector == "v2064701", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change_percent)`%) jobs in British Columbia in `r df_month_over_month_changes %>% filter(vector == "v2064701", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(date) %>% format("%B")`.

• Year-over-year to `r month(Sys.Date() %m-% months(1), label = TRUE, abbr = FALSE)` `r year(Sys.Date())`, there are `r df_year_over_year_changes %>% filter(vector == "v2064701", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(year_change) %>% comma()` (`r df_year_over_year_changes %>% filter(vector == "v2064701", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(year_change_percent)`%) fewer people employed in British Columbia.

• Youth (age 15-24) employment increased by `r df_month_over_month_changes %>% filter(vector == "v2064728", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()` (`r df_month_over_month_changes %>% filter(vector == "v2064728", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change_percent)`%) in `r month(Sys.Date() %m-% months(1), label = TRUE, abbr = FALSE)`, following a decline of `r df_month_over_month_changes %>% filter(vector == "v2064728", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 2) %>% pull(month_change) %>% comma()` (`r df_month_over_month_changes %>% filter(vector == "v2064728", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 2) %>% pull(month_change_percent)`%) in `r month(Sys.Date() %m-% months(2), label = TRUE, abbr = FALSE)`. Employment for core working group ages (25-54) increased by `r df_month_over_month_changes %>% filter(vector == "v2064836", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()` in `r month(Sys.Date() %m-% months(1), label = TRUE, abbr = FALSE)` (`r df_month_over_month_changes %>% filter(vector == "v2064836", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change_percent)`%).

• Among adults (age 25+), men’s employment increased by `r df_month_over_month_changes %>% filter(vector == "v2064818", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()` (`r df_month_over_month_changes %>% filter(vector == "v2064818", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change_percent)`%) since `r month(Sys.Date() %m-% months(2), label = TRUE, abbr = FALSE)` `r year(Sys.Date())` but decreased by `r df_year_over_year_changes %>% filter(vector == "v2064818", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(year_change) %>% comma()` since `r month(Sys.Date() %m-% months(1), label = TRUE, abbr = FALSE)` `r year(Sys.Date() %m-% years(1))` (`r df_year_over_year_changes %>% filter(vector == "v2064818", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(year_change_percent)`%). Women’s employment increased by `r df_month_over_month_changes %>% filter(vector == "v2064827", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()` (`r df_month_over_month_changes %>% filter(vector == "v2064827", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change_percent)`%) since `r month(Sys.Date() %m-% months(2), label = TRUE, abbr = FALSE)` `r year(Sys.Date())` and decreased by `r df_year_over_year_changes %>% filter(vector == "v2064827", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(year_change) %>% comma()` (`r df_year_over_year_changes %>% filter(vector == "v2064827", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(year_change_percent)`%) since `r month(Sys.Date() %m-% months(1), label = TRUE, abbr = FALSE)` `r year(Sys.Date() %m-% years(1))`.


```{r}
## Statistics Canada Table "14-10-0288-01"
## Employment by class of worker, monthly, seasonally adjusted and unadjusted, last 5 months
## https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410028801

## v2067021 (British Columbia, Both sexes, Self-employed, Seasonally adjusted, Estimate)
## v2067020 (British Columbia, Both sexes, Private-sector, Seasonally adjusted, Estimate)
## v2067019 (British Columbia, Both sexes, Public-sector, Seasonally adjusted, Estimate)


vectors2 <- c(
  "self-employed" = "v2067021",
  "private-sector" = "v2067020",
  "public-sector" = "v2067019"
)

df2 <- get_cansim_vector_for_latest_periods(vectors2,
                                           periods = 5) %>%
  normalize_cansim_values() %>%
  clean_names()

df2_month_over_month_changes <- df2 %>%
  group_by(vector) %>%
  mutate(
    month_change = value - lag(value),
    month_change_percent = round((value / lag(value) - 1) * 100,
                                 digits = 1)
  )
```

• Public (`r df2_month_over_month_changes %>% filter(vector == "v2067019", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()`) and private (`r df2_month_over_month_changes %>% filter(vector == "v2067020", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()`) sectors had increased employment in `r month(Sys.Date() %m-% months(1), label = TRUE, abbr = FALSE)`, while the number of self-employed individuals decreased by `r df2_month_over_month_changes %>% filter(vector == "v2067021", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()` (`r df2_month_over_month_changes %>% filter(vector == "v2067021", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change_percent)`%) following an increase of `r df2_month_over_month_changes %>% filter(vector == "v2067021", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 2) %>% pull(month_change) %>% comma()` in `r month(Sys.Date() %m-% months(2), label = TRUE, abbr = FALSE)`.

```{r}
## Statistics Canada Table "14-10-0127-01"
## Reason for not looking for work, monthly, unadjusted for seasonality
## https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1410012701&pickMembers%5B0%5D=1.11&pickMembers%5B1%5D=3.1&pickMembers%5B2%5D=4.1

## v54536377 (British Columbia, Total not in the labour force, Both sexes, 15 years and over)	
## v54536545 (British Columbia, Not in the labour force and did not want work or not available, Both sexes, 15 years and over)
## v54536398 (British Columbia, Not in the labour force but wanted work, Both sexes, 15 years and over)	


vectors3 <- c(
  "total-not" = "v54536377",
  "not-want-not-available" = "v54536545",
  "not-but-wanted" = "v54536398"
)

df3 <- get_cansim_vector_for_latest_periods(vectors3,
                                           periods = 5) %>%
  normalize_cansim_values() %>%
  clean_names()

df3_month_over_month_changes <- df3 %>%
  group_by(vector) %>%
  mutate(
    month_change = value - lag(value),
    month_change_percent = round((value / lag(value) - 1) * 100,
                                 digits = 1)
  )
```

• Those not in the labour force decreased in `r month(Sys.Date() %m-% months(1), label = TRUE, abbr = FALSE)` by `r df3_month_over_month_changes %>% filter(vector == "v54536377", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()`, including those who did not want to work or who were not available (`r df3_month_over_month_changes %>% filter(vector == "v54536545", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()`, or `r df3_month_over_month_changes %>% filter(vector == "v54536545", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change_percent)`%) and those who wanted work (`r df3_month_over_month_changes %>% filter(vector == "v54536398", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change) %>% comma()`, or `r df3_month_over_month_changes %>% filter(vector == "v54536398", year(date) == year(Sys.Date()), month(date) == month(Sys.Date()) - 1) %>% pull(month_change_percent)`%).





